---
title: "Correlation Practice Solutions"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, toc=TRUE)
library(tidyverse)
```

```{r}
library(tidyverse)
```

## Number of Calls and Sales
```{r message=FALSE}
SalesCalls <- read_csv("SalesCalls.csv")
attach(SalesCalls)
```

### 1 Scatterplot
```{r}
plot(sales_made ~ calls)
```

The spread of data points with no apparent pattern suggests that there is no relationship between calls and sales, that is, more calls don't lead to more sales. Based on the scatter plot one would not expect to see a moderate or strong correlation between calls and sales.

### 2 Correlation Coefficient
```{r}
cor(sales_made, calls)
```

### 3 Interpretation
The correlation coefficient is -0.099, a very weak, practically non-existent relationship between the number of calls and the number of sales made. The relationship is an inverse or negative relationship, as well, meaning that as the number of calls increase the numbers of sales made decreases. The scatter plot fully supports that interpretation.

You should tell the supervisor that there is no relationship between the number of calls made and the number of sales made. Other data is needed to understand what leads to higher sales.


## Car MPG and Weight
```{r message=FALSE}
Cars <- read_csv("Cars.csv")
attach(Cars)
```

### 1 Scatterplot
```{r}
plot(mpg ~ wt)
```

There is a pretty clear negative linear pattern here.

### 2 Correlation coefficient
```{r}
cor(mpg, wt)
```

### 3 Interpretation
The value of $r$ is close to -1, so we have a strong, negative correlation.