---
title: "Module 4 SLR Report (KEY)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

```{r}
library(tidyverse)
```

We will load the provided data.

```{r}
perf <- read_csv("PredictJobPerformance.csv")
```

## Understand the Data

First I'll check on the correlation and scatterplots of the numeric data.  The R commands below create a new data frame with just the numeric columns, then returns the correlation matrix for those columns

```{r}
summary(perf)
plot(perf)
```

The scatterplot does not look great.  The upper left is generally empty, but there is not really a strong correlation.

## Create a regression model

Let's try to build a regression model.
```{r}
mymodel <- lm(perf$Performance ~ perf$Test)
summary(mymodel)
```

The model is $Performance = 28.2897 + 0.1521 Test$.

However, the model is not statistically significant. There is a very large p-value on the F-statistic, and the $R^2$ value is quite small. This model does not show a linear relationship between test scores and performance.

Jack should not use this model to predict job performance.

