---
title: "Week 3 Quiz Solutions"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, toc=TRUE)
library(tidyverse)
```

```{r}
library(tidyverse)
```
## Levels of Measurement
Nothing to compute here, see solutions on Canvas.

## Is the new battery better?
```{r}
battery <- read_csv("Batteries.csv")
attach(battery)
```
compute the means and standard deviations:
```{r}
print("Means:")
tapply(usage_time, battery_type, mean)
print("Standard deviations:")
tapply(usage_time, battery_type, sd)

```

a boxplot to verify the ANOVA assumptions:
```{r}
boxplot(usage_time ~ battery_type)
```

Finally, the ANOVA results
```{r}
summary(aov(usage_time ~ battery_type))
```



---------------------
## Laser printer cost and speed

```{r}
printers <- read_csv("Printers.csv")
attach(printers)
```

First, the scatterplot:
```{r}
plot(Price ~ BW_pages_per_minute)
```
I see a moderate, positive, linear relationship

The correlation coefficient, $r$:
```{r}
cor(Price, BW_pages_per_minute )
```

This suggests a moderate, positive relationship.  This suggests that there are likely other characteristics that influence price.

