---
title: "Module 5 MLR Report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
library(tidyverse)
```


## Modeling Sale Price

### Question
Which characteristics best predict house sale price?

We load the data to begin.
```{r}
sales <- read_csv("kc_HouseSales.csv")
attach(sales)
```

### Understand the Data

First I'll check on the correlation of the numeric data.  We'll use the `cor` command to find correlation to `price` for all but the `date` column (since that is not numerical).  Since `date` is the second column, we can exclude it with this command.

```{r}
cor(sales[, -2], price)
```

Looking at correlation with `price` it seems most of these variables have at least a moderate correlation except for `id`, `sqft_lot`, `condition`, `yr_built`, and `zipcode`.  

Let's try to make a model with all the other variables.
```{r}
summary(lm(price ~ bedrooms + bathrooms + sqft_living + floors + waterfront + view + grade + sqft_above + sqft_basement + yr_renovated + lat))
```

Notice the NA for `sqft_basement`. That is `R` indicating that variable is too closely related to the others. So we don't need it in our model.
```{r}
summary(lm(price ~ bedrooms + bathrooms + sqft_living + floors + waterfront + view + grade + sqft_above + yr_renovated + lat))
```

Now it looks like the least significant variable is `sqft_above`.
```{r}
summary(lm(price ~ bedrooms + bathrooms + sqft_living + floors + waterfront + view + grade + yr_renovated + lat))
```

Now we will need to remove `bathrooms`.
```{r}
summary(lm(price ~ bedrooms + sqft_living + floors + waterfront + view + grade + yr_renovated + lat))
```

This model looks good so far. It is significant over all, each variable is significant, and the $R^2$ value is 0.66 indicating these variables explain 66% of the variation in sale price.
