---
title: "Module 5 MLR Practice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

```{r}
library(tidyverse)
```

## Cafe Data
```{r}
cafe <- read_csv("CafeData2.csv")
attach(cafe)
```


I'll need to create a new column for total coffee sales.  Since the problem asked for all coffee drinks I'll add the four relevant columns and create a new column called `total_coffee`.
```{r}
cafe$total_coffee <- `Coffee (Large 16oz)`+`Coffee (Medium 12oz)`+Lattes+Mochas
attach(cafe)
```

Now `DayOfWeek` is a nominal variable, so we expect `R` to create dummy variables for it when creating linear models.

### Total Coffee vs Day of Week

Let's try a linear model with the day of the week predicting coffee sales.
```{r}
summary(lm(total_coffee ~ DayOfWeek))
```

This does not look very good. The model is only barely significant (p-value 0.03), and the $R^2$ value is only 0.11.

### Total Sales vs Day of Week

Now we will see if day of the week can explain `TotalSales`
```{r}
summary(lm(TotalSales ~ DayOfWeek))
```

This model is very promising.  It is statistically significant, and the $R^2$ is 0.77.

The equation is
$$TotalSales = 134.19 + 113.56 * Mon + 298.62 * Tue + 293.04 * Wed + 305.83 * Thu$$

From this we can predict average sales on a Monday will be $134.19+113.56 = 247.75$.

We predict average sales on Fridays to be $134.19$.  


### Summary
It looks like the day of the week can help to predict `TotalSales` but not necessarily total coffee drinks.




