---
title: "Week 7 Plugging Values into a Model"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

## Transformed Predictor Variables

For this sample report, I'll use a portion of `R`'s built-in data set `mtcars`. I created a data file with five of the variables from that set for the purposes of this sample report.

We load the data set.
```{r, message=FALSE}
library(tidyverse)
```

```{r}
myData <- read_csv("carmpg.csv")
attach(myData)
head(myData)
```

The variable `gear` only has three values which seems to suggest that we should treat it as a categorical variable, not numerical. We can tell `R` to do this by specifying that `gear` is a factor. This will tell `R` to create dummy variables when making the model.
```{r}
nominal_gear <- factor(gear)
```

When checking correlations, I found displacement and horsepower both show some non-linearity. We will transform both and then attempt to build the model.

```{r}
ln_disp <- log(disp)
ln_hp <- log(hp)
```

You can verify that the new scatterplots show a much more linear relationship and the correlation coefficients for `mpg` with both `ln_disp` and `ln_hp` have increased.

## Building the Model

Let's construct a model and attempt to make a prediction from it. For the purposes of this exercise, I'm not going to worry about which variables are significant.
```{r}
mpgModel <- lm( mpg ~ ln_disp + ln_hp + wt + nominal_gear)
mpgModel
```

Now let's predict the mpg for a car with displacement 200, 150 horsepower, weight 3.5, and 4 gears. Be careful when plugging in since some of these are transformed. In particular, note the quotes around the nominal variable.
```{r}
values <- data.frame( ln_disp = log(200), ln_hp = log(150), wt=3.5, nominal_gear = "4")
predict(mpgModel, values, interval="predict")
```

## Other transformations

What if we've created a model where we transformed the response variable?
With the Mammal Gestation data in [Example 9-2](https://newonlinecourses.science.psu.edu/stat501/lesson/9/9.2#paragraph--756) of [Chapter 9 in the Penn State e-book](https://newonlinecourses.science.psu.edu/stat501/lesson/9) on regression, the authors conclude the only problem with the data is unequal variances in the errors. They decide to transform the response variable to address this problem.

Let's load the data, apply a `log` transformation to the response variable, then generate the new model.
```{r message=FALSE}
gest <- read_csv("mammgest.csv")
attach(gest)
```

```{r}
ln_Gestation <- log(Gestation)
gest_NEWmodel <- lm( ln_Gestation ~ Birthwgt)
gest_NEWmodel
```


Let's make a prediction for a birthweight of 45.
```{r}
predict( gest_NEWmodel, data.frame( Birthwgt = 45) )
```

Now is this the prediction for gestation length?  No, its the prediction for `ln_Gestation`

```{r}
exp(5.747278)
```

How about a prediction interval?
```{r}
predict( gest_NEWmodel, data.frame( Birthwgt = 45), interval = "predict" )
```

```{r}
exp(5.234582)
exp(6.259974)
```

So our prediction interval is between 187.6507 and 523.2053.
